---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

extrafont::loadfonts(device = "win")
```

# dunnr

<!-- badges: start -->
<!-- badges: end -->

The goal of `dunnr` is to compile various helper functions and lists that I frequently use.
Mostly, it has functions for plotting in `ggplot2`.

## Installation

``` r
# install.packages("remotes")
remotes::install_github("taylordunn/dunnr")
```
## Examples

### Plotting functions

My go-to theme can be applied with `theme_td()`:

```{r example_theme_td}
library(dunnr)
library(ggplot2)
library(dplyr)
library(palmerpenguins)

p1 <- penguins %>%
  filter(!is.na(bill_length_mm)) %>%
  ggplot(aes(x = bill_length_mm, y = bill_depth_mm, color = species)) +
  geom_point(aes(shape = species), size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
  scale_color_brewer(palette = "Set1") +
  labs(title = "Penguin bill dimensions",
       subtitle = "Bill length and depth for different penguin species",
       x = "Bill length (mm)", y = "Bill depth (mm)",
       color = "Penguin species", shape = "Penguin species",
       caption = "Data from the palmerpenguins package") +
  facet_wrap(~species, nrow = 1)
p1 + theme_td()
```

Note that the default font family here is Roboto Condensed, which can be downloaded from [Google Fonts](https://fonts.google.com/specimen/Roboto+Condensed) (to save headaches on Windows, use the "Install for all users" option so that R can find the font easily).
Then to import and load the font:

```{r eval=F}
extrafont::font_import(pattern = "Roboto", prompt = FALSE)
extrafont::loadfonts(device = "win")
```

Alternatively, give another font as an argument, e.g. `theme_td(base_family = "Arial")`.

### `add_facet_border()`

Sometimes when plotting facets like above, it helps to have some visual separation in the panels.
The `add_facet_border()` function is a quick way to do so without adjusting `theme()` elements directly:

```{r example_add_facet_borders}
p1 + theme_td() + add_facet_borders()
```

### `set_geom_fonts()`

The `set_geom_fonts()` function sets the default fonts of `geom_text()` and `geom_label()`.
If no font is specified, it will automatically take the font from the current `ggplot2` theme:

```{r example_set_geom_fonts, fig.height=3, fig.width=6}
library(glue)
p <- penguins %>%
  count(species, island, name = "n_penguins") %>%
  ggplot(aes(y = species, x = n_penguins)) +
  geom_col(aes(fill = island)) +
  geom_label(aes(label = glue("{island}: n = {n_penguins}")),
             position = "stack", hjust = 1) +
  scale_fill_brewer(palette = "Set1") +
  scale_x_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(title = "Number of penguins per island",
       subtitle = "Different font",
       x = "Number of penguins", y = NULL,
       caption = "Data from the palmerpenguins package") +
  theme(legend.position = "none")

# By default, geom_label() doesn't have the same font as theme_td()
p + theme_td() +
  theme(legend.position = "none")

# But if we set the theme
theme_set(theme_td())
# And then set fonts with the helper function
set_geom_fonts()

p + theme_td() +
  theme(legend.position = "none") +
  labs(subtitle = "Same font")
```

